---
title: "info550_project"
author: "Carol Liu"
date: "10/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

pkgs <- c("ggplot2", "reshape2","gridExtra","kableExtra")
for(p in pkgs){
	if(!(p %in% pkgs)){
		install.packages(p)
	}
}

library(ggplot2)
library(reshape2)
library(gridExtra)
library(kableExtra)

contacts<-readRDS("usa_contacts")
georgiapop<-read.csv("georgiapop.csv")
parms <- read.csv("parms.csv")
int_scen <- read.csv("intervention_scenarios.csv")

```

# Background

As mentioned in the README.md, we use use a simple SEIR-like model I previously worked on to simulate cases of severe acute respiratory syndrome coronoavirus 2 (SARS-CoV2) in Georgia. State-level shelter-in-place orders began on March 16th where schools, work places and leisure areas were closed. At the time, I was subsequently interested in the impact of different social distancing strategies on infection transmission. The goal of my model was to address the following primary research question (from early May): how will the epidemic differ if current shelter-in-place orders are relaxed with a 50% return to work and no additional intervention versus a 100% return to work and either 1) additional measures to social distance at home; 2) isolation of 20% of pre-infectious cases identified from contact tracing or 3) isolation of 50% of pre-infectious cases identified from contact tracing?

The methods, model structure, input parameters and intervention scenarios are described in the README.md as well.

```{r int_matrices}

## list contact-based intervention scenarios

interventions<-  list(
  # constraints under a DO-NOTHING scenario 
  base =list(home = diag(1,16,16),
             work = diag(1,16,16),
             school = diag(1,16,16),
             others = diag(1,16,16)),
  # Georgia lockdown--assume from XX Jan to XX Feb 
  shelterinplace = list(home = diag(1,16,16),
                        work = diag(0.1,16,16),
                        school = diag(0,16,16),
                        others = diag(c(rep(0.1,4),rep(0.25,12)))),
  # constraints under school closure but 80% workplace
  relax = list(home = diag(c(rep(1,4),rep(1,12))),
                      work = diag(0.5,16,16),
                      school = diag(0,16,16),
                      others = diag(c(rep(0.5,4),rep(0.5,12)))), 
  # constraints under household distancing and school closure known infected cases only
  relaxhomedist = list(home = diag(0.5,16,16),
                      work = diag(1,16,16),
                      school = diag(0,16,16),
                      others = diag(0.5,16,16)))

```


```{r SEIR model}
# Children as infectious and as susceptible 
SEIR = function(prop.isolate = 0, hosp.rate =0.0075,
                dateShelterInPlace = as.Date('2021-12-31') , 
                dateRelaxIntervention = as.Date('2021-12-31'),
                dateRelaxandHomedistance = as.Date('2021-12-31'),
                dateStart = as.Date('2020-02-01'),POP = georgiapop,durInf = 7,contact_ga=contacts)
{
  
  # Load population information
  pop = list()
  pop$N = sum(POP$popage)
  pop$p_age = georgiapop$propage
  N_age = pop$N*pop$p_age                                        # Population age structure (in numbers)
  
  # Specify epi info
  durLat = 5.2;   	                                             # Mean latent period (days) from Backer, et al (2020)
  gamma = 1/durInf;                                              # removal rate
  alpha = 1/durLat;                                      
  tau1 = prop.isolate
  hosp.rate = hosp.rate
  dt = 1;                                                        # Time step (days)
  tmax = 400;                                                    # Time horizon (days) 366 days in 2020 cause of leap year
  numSteps = tmax/dt;  	                                         # Total number of simulation time steps
  dateStart = as.Date(dateStart)                                 # included as a function argument 
  dateEnd = dateStart+(tmax-1)
  
  # Declare the state variables and related variables:
  S = E = Ei = I = R = H = array(0,c(numSteps,length(pop$p_age)))
  lambda = infections = infectious = incidence = isolate = newhosp = reported = cumulativeIncidence = array(0,c(numSteps,length(pop$p_age)))
  time = array(0,numSteps)
  
  # Initialise the time-dependent variables, i.e. setting the values of the variables at time 0
  E[1,] = 0
  Ei[1,]=0
  I[1,] = rep(1,16) 
  H[1,] = 0 
  R[1,] = 0 
  S[1,] = N_age-E[1,]-Ei[1,]-I[1,]-H[1,]-R[1,]
  newhosp[1,] = 0;
  incidence[1,] = 0;
  isolate[1,] = 0; 
  reported[1,] = 0;
  time[1] = 0;
  
  ## Time of interventions
  tShelterInPlace = as.vector(dateShelterInPlace - dateStart)+1
  tRelaxIntervention = as.vector(dateRelaxIntervention - dateStart)+1
  tRelaxandHomedistance = as.vector(dateRelaxandHomedistance - dateStart)+1

  for (stepIndex in 1: (numSteps-1))
  { 
    # load plausible intervetions 
    constraintsIntervention = interventions
    # I0: before shelter in place, use base contact rates
    if(time[stepIndex] < tShelterInPlace)  
    {
      CONSTRAINT = constraintsIntervention$base
    } else if(time[stepIndex] >= tShelterInPlace & time[stepIndex] < tRelaxIntervention) 
    # I1:  when shelter in place starts use 'shelterinplace' contact scenarios
     {
        INTERVENTION = "shelterinplace"   
        CONSTRAINT = constraintsIntervention[[INTERVENTION]] 
     } else if(time[stepIndex] >= tRelaxIntervention & time[stepIndex] < tRelaxandHomedistance)
    # I2:  When interventions relax, use 'relax' contact scenarios
     {
        INTERVENTION = "relax"   
        CONSTRAINT = constraintsIntervention[[INTERVENTION]] 
     }  else if(time[stepIndex] >= tRelaxandHomedistance)
       # I2:  If relaxing is with home distancing, then use "relaxhomedist". 
     {
       INTERVENTION = "relaxhomedist"   
       CONSTRAINT = constraintsIntervention[[INTERVENTION]] 
     } 
    
    C = CONSTRAINT[[1]]%*%contact_ga[[1]]+
        CONSTRAINT[[2]]%*%contact_ga[[2]]+
        CONSTRAINT[[3]]%*%contact_ga[[3]]+
        CONSTRAINT[[4]]%*%contact_ga[[4]]
    
    lambda[stepIndex,] = as.numeric(0.035)*(as.matrix(C)%*%as.matrix(I[stepIndex,]/N_age))
    
    tau <- ifelse(stepIndex < 90, 0,tau1)                #Date switch for proportion isolated
    
    
    numInfection = lambda[stepIndex,]*S[stepIndex,]*dt;   #S to E
    numInfectious = alpha*(1-tau)*E[stepIndex,]*dt;       #E to I
    numIsolate = alpha*tau*E[stepIndex,]*dt;              #E to Ei
    numNewHosp = hosp.rate*I[stepIndex,];                 #I to H
    numRecInfectious = gamma*I[stepIndex,]*dt;            #I to R
    numRecIsolate = gamma*Ei[stepIndex,]*dt;              #Ei to R
    numRecHosp = gamma*H[stepIndex,]*dt;                  #H to R

    # SEIR difference equations 
    S[stepIndex+1,] = S[stepIndex,]-numInfection;
    E[stepIndex+1,] = E[stepIndex,]+numInfection-numInfectious-numIsolate;
    Ei[stepIndex+1,] = Ei[stepIndex,]+numIsolate-numRecIsolate;
    I[stepIndex+1,] = I[stepIndex,]+numInfectious-numRecInfectious - numNewHosp;
    H[stepIndex+1,] = H[stepIndex,]+numNewHosp - numRecHosp ;
    R[stepIndex+1,] = R[stepIndex,]+numRecInfectious + numRecIsolate + numRecHosp;
    infections[stepIndex+1,] = numInfection/dt;
    infectious[stepIndex+1,] = numInfectious/dt;
    isolate[stepIndex+1,] = numIsolate/dt;
    incidence[stepIndex+1,] = (numInfectious+numIsolate)/dt;
    newhosp[stepIndex+1,] = numNewHosp/dt;
    time[stepIndex+1] = time[stepIndex]+dt;
    
  }
  output = list(S =S, E = E, Ei = Ei, I = I, R = R, time = time, lambda=lambda,
                infections = infections, infectious = infectious, incidence = incidence, isolate = isolate, newhosp=newhosp,
                N_age= N_age, 
                dateStart = dateStart, dateEnd = dateEnd, C=C,beta=beta, tShelterInPlace=tShelterInPlace, tRelaxIntervention=tRelaxIntervention)
  return(output)
}
```

# Results
Below is a plot of new infections over time under the six intervention scenarios for all age groups. 
```{r simulations}
#################
###Simulations###
#################

epi_donothing = SEIR(dateShelterInPlace = as.Date('2021-11-01'),
                                    dateRelaxIntervention = as.Date('2021-12-31'))

epi_shelterinplace = SEIR(dateShelterInPlace = as.Date('2020-03-16'),
                                    dateRelaxIntervention = as.Date('2021-12-31'))

epi_shelterrelax = SEIR(dateShelterInPlace = as.Date('2020-03-16'),
                                    dateRelaxIntervention = as.Date('2020-05-01'))

epi_relaxhomedist = SEIR(dateShelterInPlace = as.Date('2020-03-16'),
                                    dateRelaxIntervention = as.Date('2020-04-30'),
                                    dateRelaxandHomedistance = as.Date('2020-05-01'))

epi_relaxisolate1 = SEIR(prop.isolate = 0.2, dateShelterInPlace = as.Date('2020-03-16'),
                         dateRelaxIntervention = as.Date('2020-05-01'))

epi_relaxisolate2 = SEIR(prop.isolate = 0.5, dateShelterInPlace = as.Date('2020-03-16'),
                        dateRelaxIntervention = as.Date('2020-05-01'))

######################################
#####Summaries and Plots below#######
#####################################

## Total incidence
date<-seq(from = as.Date('2020-02-01'), to = as.Date('2020-02-01') + 399, by = 1)
epi_totincidence<-data.frame(time=date,Donothing=rep(0,400),IndefiniteShelter=rep(0,400),ShelterRelaxWork=rep(0,400),ShelterRelaxHomedist=rep(0,400),
                             RelaxIsolate1=rep(0,400),RelaxIsolate2=rep(0,400))

epi_totincidence["Donothing"]<-rowSums(epi_donothing$incidence)
epi_totincidence["IndefiniteShelter"]<-rowSums(epi_shelterinplace$incidence)
epi_totincidence["ShelterRelaxWork"]<-rowSums(epi_shelterrelax$incidence)
epi_totincidence["ShelterRelaxHomedist"]<-rowSums(epi_relaxhomedist$incidence)
epi_totincidence["RelaxIsolate1"]<-rowSums(epi_relaxisolate1$incidence)
epi_totincidence["RelaxIsolate2"]<-rowSums(epi_relaxisolate2$incidence)

epi_totincidence<-melt(epi_totincidence,id="time")
  
p1<-ggplot(epi_totincidence ,aes(x=time,y=value))+
  geom_line(size=1,aes(x=time,y=value,color=variable))+
  #scale_color_manual(values=c("red","black"),labels=c("Projected", "Observed"))+
  xlab("Date")+ylab("New infections") + 
  theme_minimal()+
  #geom_rect(aes(xmin=as.Date('2020-03-16'), xmax=as.Date('2020-05-01'), ymin=0, ymax=Inf))
  geom_vline(xintercept=as.Date('2020-03-16'), linetype=1) +
  theme(legend.position="bottom",legend.title=element_blank())+
  ggtitle("New infections over time in Georgia under six intervention scenarios for all age-groups")

p1
```

# SUmmary results

Below is some summary statistics under the six interventions. Specifically, we explore the peak date of incidence, number of new infections at peak incidence and the total infections from until the end of the simulation (spring 2021). We see that in a do nothing scenario, the cumulative cases is simulated to be `r round(sum(rowSums(epi_donothing[["incidence"]])), digits =0)`, in a scenario of indefinite shelter-in-place and perpetual maximum social distancing, the cumulative cases is simulated to be `r  round(sum(rowSums(epi_shelterinplace[["incidence"]])), digits =0)`. In reality, Georgia currently has over 300,000 cases. Alas, all models are wrong, some are useful. 

``` {r summary_results}
######################################
#######Summary results     #################
######################################

results <- data.frame(Intervention = c("DoNothing","IndefiniteShelter",
                                       "ShelterRelaxWork","ShelterRelaxHomeDist",
                                       "RelaxIsolate1","RelaxIsolate2"),
                      PeakDate = c(
                       as.Date('2020-02-01')+which.max(rowSums(epi_donothing[["incidence"]])),
                      as.Date('2020-02-01')+which.max(rowSums(epi_shelterinplace[["incidence"]])),
                      as.Date('2020-02-01')+which.max(rowSums(epi_shelterrelax[["incidence"]])),
                      as.Date('2020-02-01')+which.max(rowSums(epi_relaxhomedist[["incidence"]])),
                      as.Date('2020-02-01')+which.max(rowSums(epi_relaxisolate1[["incidence"]])),
                      as.Date('2020-02-01')+which.max(rowSums(epi_relaxisolate2[["incidence"]]))),
                      
                      PeakIncidence = c(
                      round(max(rowSums(epi_donothing[["incidence"]])), digits=0),
                      round(max(rowSums(epi_shelterinplace[["incidence"]])), digits=0),
                      round(max(rowSums(epi_shelterrelax[["incidence"]])),digits=0),
                      round(max(rowSums(epi_relaxhomedist[["incidence"]])), digits=0),
                      round(max(rowSums(epi_relaxisolate1[["incidence"]])), digits=0),
                      round(max(rowSums(epi_relaxisolate2[["incidence"]])),digits=0)),

                      Totalnfections = c(
                      round(sum(rowSums(epi_donothing[["incidence"]])), digits =0),
                      round(sum(rowSums(epi_shelterinplace[["incidence"]])), digits =0),
                      round(sum(rowSums(epi_shelterrelax[["incidence"]])), digits=0),
                      round(sum(rowSums(epi_relaxhomedist[["incidence"]])), digits = 0),
                      round(sum(rowSums(epi_relaxisolate1[["incidence"]])), digits =0),
                      round(sum(rowSums(epi_relaxisolate2[["incidence"]])), digits =0)))

kable(results)

```




